---
title: "Untitled"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      fig.width = 9)

library(tidyverse)
library(lubridate)
library(RColorBrewer)
```

Here we try to vizualize simple seasonal chart using ggplot2. This will be used for making series of charts by trelliscopjs 

#### 1) Load data
```{r}
water_data <- readRDS("data/processing/clean_water_data.Rds")
```
##### 2) Obzerve data
```{r}
glimpse(water_data)
```
#### 3) Make dataset for representing monthly changes by water parameters via years 2008-2017 for each water region parameter
```{r}
df_season <- water_data %>% 
  select(-(1:3), -(5:7)) %>%
  mutate(Year = factor(year(Data_sposterezhen), ordered = T, levels = seq(2017, 2008, -1)),
         Month = month(Data_sposterezhen, label = T)) %>% 
  dplyr::filter(Data_sposterezhen < "2018-01-01") %>% 
  select(1, 2, Year, Month, everything()) %>% 
  gather(key = Parameter, value = Value, -(1:4)) %>% 
  select(-Data_sposterezhen) %>% 
  group_by(Parameter, Rajon_richkovoho_basejnu, Year, Month) %>% 
  summarise(Value = mean(Value))

glimpse(df_season)
```
#### 4) Visualize one parameter for one water region
```{r}
df_season %>% 
  dplyr::filter(Parameter == 'Kysen_rozchynenyj_MhO2_per_dm3',
                Rajon_richkovoho_basejnu == 'Дунай') %>%
  ggplot(aes(x = Month, y = Value)) +
    geom_area(aes(fill = Year, group = Year), color = "white", alpha = 0.8) +
    scale_x_discrete(expand = c(0.01, 0.01)) +
    scale_color_manual(values = brewer.pal(10, "Set1")) +
    labs(title = "Зміна середньомісячного значення Розчиненого кисню впродовж кожного року",
         x = NULL) +
    theme_classic() +
    theme(panel.grid.major.x = element_line(colour="black"),
          axis.text = element_text(size = 12, colour = "black"),
          legend.position = "right",
          legend.title = element_text(face = "bold", size = 13),
          legend.text = element_text(size = 12))
```
Зауваження! На графіку важливими є положення шарів (ліній розмежування) відносно один одного - не їх абсолютні значення.
#### 5) Visualize one parameter for all water regions
```{r}
df_season %>% 
  dplyr::filter(Parameter == 'Kysen_rozchynenyj_MhO2_per_dm3') %>%
  ggplot(aes(x = Month, y = Value)) +
    geom_area(aes(fill = Year, group = Year), color = "white", alpha = 0.8) +
    facet_wrap(~ Rajon_richkovoho_basejnu, nrow = 4, ncol = 2, scales = "free_y") +
    scale_color_manual(values = brewer.pal(10, "Set1")) +
    labs(title = "Зміна середньомісячного значення Розчиненого кисню впродовж кожного року",
         x = NULL, y = NULL) +
    theme_classic() +
    theme(
      panel.grid.major.x = element_line(colour="black")
    )
```
